---
title: "LincolnLoC"
format: pdf
editor: visual
---

## Text Analysis of Lincoln Speeches and Correspondences (1843-1858) and (1863-1865)

The following Quarto document conducts an analysis on the word frequency and lexical diversity of Lincoln's speeches and correspondences approximately around Frederick Douglass's autobiographies *Narrative Life of Frederick Douglass* (1845) and *My Bondage and My Freedom* (1855), as well as the time around the Emancipation Proclamation (1863) and end of life. The purpose of these test are to consider our evaluation methods in both their ability to transfer to another work outside of Douglass's and provide additional historical context for Douglass's writing, with the potential to be further explored later.

The text file for this project was pulled from Project Gutenburg's eBook "The Papers and Writing of Abraham Lincoln" and uses two specific volumes: Volume 2 (1843-1858) and Volume 7 (1863-1865).

The Papers and Writings of Abraham Lincoln, Volume 2 was released on Project Gutenberg as eBook #2654. The Papers and Writings of Abraham Lincoln, Volume 7 was released as eBook #2659. Both volumes were released in September 30, 2004 and were used as the version released in October 29, 2012.

## Calculating Word Frequencies of "The Papers and Writing of Abraham Lincoln"

The word frequency of Abraham Lincoln's writings will be tested to Douglass's first two biographies ("Narrative Life of Frederick Douglass" (1845) and "My Bondage and My Freedom" (1855)). Since "Life and Times of Frederick Douglass" (1881) and Douglass's travel diary (1886-1887) were written after Lincoln's death, these works are not considered.

```{r}
library(gutenbergr)
library(tidyverse)
library(tidytext)
library(dplyr)

# retrieving the documents from project gutenberg
PWAL2.v <- gutenberg_download(2654)
tidy_PWAL2 <- PWAL2.v %>%
  unnest_tokens(word, text) %>%
  anti_join(stop_words) %>%
  count(word, sort = TRUE) %>%
  rename(word_count = n)
mean_wf <- mean(tidy_PWAL2$word_count)

PWAL7.v <- gutenberg_download(2659)
tidy_PWAL7 <- PWAL7.v %>%
  unnest_tokens(word, text) %>%
  anti_join(stop_words) %>%
  count(word, sort = TRUE) %>%
  rename(word_count = n)
mean_wf <- mean(tidy_PWAL2$word_count)
```

```{r}
# counting the word frequencies from Abraham Lincoln's writings
frequencyAL <- bind_rows(mutate(tidy_PWAL2, work="Volume 2 (1843-1858)"),
                       mutate(tidy_PWAL7, work="Volume 7 (1863-1865)"))
frequencyAL <- frequencyAL %>% spread(work, word_count, fill=0)

# loading the files for Douglass's word frequencies
frequencyFD <- read_csv("Word_Count_Frequencies.csv", show_col_types = FALSE)
frequencyFD <- select(frequencyFD,word, "My Bondage (1855)", "Narrative Life (1845)")

# combining the two frequencies to better compare works
frequencyALFD <- merge.data.frame(frequencyAL, frequencyFD, by = "word")

# calculating the sum of the words
frequencyAL <- frequencyAL %>% 
  mutate(sum_word_counts= rowSums(across(c("Volume 2 (1843-1858)", "Volume 7 (1863-1865)"))))
frequencyALFD <- frequencyALFD %>% 
  mutate(sum_word_counts= rowSums(across(c("Volume 2 (1843-1858)", "Volume 7 (1863-1865)", "My Bondage (1855)", "Narrative Life (1845)"))))
         
# reordering the data by total count of word frequency
frequencyAL <- frequencyAL %>% arrange(desc(sum_word_counts))
frequencyALFD <- frequencyALFD %>% arrange(desc(sum_word_counts))

# exporting the frequency as a csv file
write.csv(frequencyAL, file="Word_Count_Frequencies_Lincoln.csv")
```
# Visualising a comparision of Lincoln's work to Douglass's Autobiographies
For the comparison of Lincoln's work to Douglass's Autobiographies, Lincoln's writings from Volume 2 (1843-1858) was used as the baseline. 
```{r}
# calculating the normalized distance of word frequencies
norm_dist <- function(work1, vol2, sum) (work1-vol2)/sum #normalized distance formula
norm_freq_ALFD <- frequencyALFD %>%
    mutate("Narrative Life (1845)" = norm_dist(frequencyALFD$"Narrative Life (1845)", 
                                               frequencyALFD$"Volume 2 (1843-1858)", 
                                               sum_word_counts)) %>%
    mutate("My Bondage (1855)" = norm_dist(frequencyALFD$"My Bondage (1855)", 
                                           frequencyALFD$"Volume 2 (1843-1858)", 
                                           sum_word_counts)) %>%
    mutate("Volume 2 (1843-1858)" = norm_dist(frequencyALFD$"Volume 2 (1843-1858)", 
                                      frequencyALFD$"Volume 2 (1843-1858)", 
                                      sum_word_counts)) %>%
    mutate("Volume 7 (1863-1865)" = NULL)

# creating a subset of the data to exclude words under a cutoff
cutoff <- 15
norm_long_ALFD <- norm_freq_ALFD %>% pivot_longer(cols = -c(word, 
                                                            sum_word_counts,
                                                            "Volume 2 (1843-1858)"), 
                                                  names_to = "Works", 
                                                  values_to = "distance")
norm_subset_ALFD <- norm_long_ALFD[norm_long_ALFD$sum_word_counts > cutoff,]

# visualising the data
ggplot(norm_subset_ALFD, aes(x = word)) +
  geom_point(aes(y= distance, color= Works, group= Works, alpha= .0005)) +
  geom_point(aes(y= norm_subset_ALFD$"Volume 2 (1843-1858)", color= "Volume 2 (1843-1858)")) +
  facet_wrap(vars(Works)) +
  guides(alpha = FALSE) +
  labs(x= "Words", y= "Similarity to Lincoln's Writings (1843-1858)") +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

```

# Analysis of Hapax of Lincoln's works
Note that our research uses the Volume division as project gutenburg to as a guide to dividing two different era's in Lincoln's life, within the context of Frederick Douglass's life.
```{r}
# calculating the lengths of the writing
total_words <- function(text) {
  words <- strsplit(text, "\\s+")[[1]]
  return(length(words))
}

total_words_PWAL2 <- sum(sapply(PWAL2.v$text, total_words))
total_words_PWAL7 <- sum(sapply(PWAL7.v$text, total_words))

# calculating hapax ACROSS ALL WORKS and then also gives the sum of hapax words in all works
hapax_words_AL <- frequencyAL[frequencyAL$sum_word_counts == 1,]
hapax_sum_AL <- colSums(hapax_words_AL[,!names(hapax_words_AL) %in% c('word', "sum_word_counts")]) #returns Vol2, Vol7, sum_words
hapax_percentage_all_AL <- unname(hapax_sum_AL)/(total_words_PWAL2+total_words_PWAL7)

# calculating hapax across individual volumes

# visualising hapax percentage
hapax_data_AL <- data.frame(Book = c("Vol.2", "Vol.7"),
                         Hapax_Percentage_AL = hapax_percentage_all_AL)
ggplot(hapax_data_AL, aes(x = Book, y = Hapax_Percentage_AL)) +
  geom_bar(stat = "identity", fill = "skyblue", color = "black") +
  geom_text(aes(label = sprintf("%.3f%%", Hapax_Percentage_AL)),
            vjust = -0.5, size = 3) + # Add percentage labels
  labs(title = "Percentage of Hapax Words in Lincoln's Writings",
       x = "Volumes", y = "Percentage of Hapax Words") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```
